---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

The base dataset is in zip file. Ensure we have it extracted and read it into data frame. Aslo setup some default knitr options.

```{r}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

dataset.zipfile <- file.path(".", "activity.zip")
dataset.csvfile <- file.path(".", "activity.csv")

if (!file.exists(dataset.csvfile)) {
    unzip(dataset.zipfile)
}
activity.data <- read.csv(dataset.csvfile)
```

## What is mean total number of steps taken per day?

Using dplyr for grouping and aggregation. Tranform date from factor to Posixlt type. Using ggplot contruct histogram of steps by date. Finally display total, mean and median by date.

```{r}
library(dplyr)
library(ggplot2)
library(scales)

total_steps_by_date <- activity.data %>% 
    group_by(date) %>%
    summarise(Count = n(), Total = sum(steps), 
              Mean = mean(steps, na.rm = TRUE), 
              Median = median(steps, na.rm = TRUE)) %>%
    transform(Date = strptime(date, "%Y-%m-%d")) %>%
    select(Date, Count, Total, Mean, Median)

p <- ggplot(total_steps_by_date, aes(x=Date, y=Total))
p + geom_histogram(stat="identity") +
    labs(title = "Total Steps By Day", x = "Day", y = "Total Steps")

total_steps_by_date
```

## What is the average daily activity pattern?

Compute means for each interval across all days. Plot this, and display the interval with the maximum average number of steps.

```{r}
interval_means <- activity.data %>%
    group_by(interval) %>%
    summarise(Mean = round(mean(steps, na.rm = TRUE), 0), Median = median(steps, na.rm = TRUE)) %>%
    select(Interval = interval, Mean, Median)

pm <- ggplot(interval_means, aes(x=Interval, y=Mean))
pm + geom_line() + labs(title = "Average Steps By Interval", y = "Avg Steps")

filter(interval_means, Mean == max(interval_means$Mean))
```

## Imputing missing values

The number of rows with missing data are:

```{r}
nrow(filter(activity.data, is.na(steps)))
```

Impute values using the mean for the corresponding interval and merge into original data set. Again plot historgram and table with Total, Mean and Median by date.

```{r}
imputed_activity_data <- select(interval_means, Interval, IntervalMean = Mean) %>%
    inner_join(activity.data, by = c("Interval" = "interval")) %>%
    mutate(steps = ifelse(is.na(steps), IntervalMean, steps)) %>%
    select(Interval, Steps = steps, Date = date)

imputed_total_steps_by_date <- imputed_activity_data %>% 
    group_by(Date) %>%
    summarise(Count = n(), Total = sum(Steps), 
              Mean = mean(Steps, na.rm = TRUE), 
              Median = median(Steps, na.rm = TRUE)) %>%
    transform(Date = strptime(Date, "%Y-%m-%d"))

ph_imputed <- ggplot(imputed_total_steps_by_date, aes(x=Date, y=Total))
ph_imputed + geom_histogram(stat="identity") +
    labs(title = "Total Steps By Day (Imputed Values)", x = "Day", y = "Total Steps")

imputed_total_steps_by_date
```

## Are there differences in activity patterns between weekdays and weekends?

Using imputed activity data, construct a factor for weekend/weekday, note we don't carry through the PosixDate field as it is unsupported by dplyr.

Then group the data by Weekday and Interval and plot it, using facet.

```{r}
activity_data_with_weekday <- imputed_activity_data %>%
    transform(PosixDate = strptime(Date, "%Y-%m-%d")) %>%
    mutate(Weekday = factor(!grepl("Sat|Sun", weekdays(PosixDate, abbreviate = TRUE)), labels = c("weekend", "weekday"))) %>%
    select(Date, Interval, Steps, Weekday)

interval_means_with_weekday <- activity_data_with_weekday %>%
    group_by(Weekday, Interval) %>%
    summarise(Mean = round(mean(Steps, 0))) %>%
    select(Interval, Mean, Weekday)

p_by_weekday <- ggplot(interval_means_with_weekday, aes(x=Interval, y=Mean))
p_by_weekday + geom_line() + facet_grid(. ~ Weekday) + labs(title = "Average Steps By Interval", y = "Avg Steps")
```
